Word learning at the semantics pragmatics interface

### 2020-04-14 paper/bda_fig.tex
update supplement working in comments from MH

    \end{tabular}
  \end{center}
  \caption{Graphical model representing the Bayesian data analysis for \emph{Integration model} used for \emph{Explanation}. $L_1$ represents the RSA model defined by Eq. X, used to predict the Combination Data from Expt.~3 $d^{CM}$ for each unique age bin $i$, lexical item $j$, and alignment condition $k$. This RSA model takes as input a speaker optimality parameter $\alpha_i$ (which varies by age), a semantic knowledge parameter $\theta_{ij}$, and a common ground sensitivity $\rho_i$ (which varies by age). Speaker optimality and semantic knowledge parameters are additionally constrained by the data from the Mutual Exclusivity task ($d^{ME}$; Expt.~1), via a related RSA model $L'_1$;  common ground sensitivity parameters $\rho_i$ are directly constrained by the data from the Common Ground experiment ($d^{CG}$; Expt.~2).  
  Each of these RSA parameters is sampled from a linear or logistic regression model that track developmental change (with intercepts and slopes given by $\beta_0$ and $\beta_1$, respectively). Additionally, the semantic knowledge parameters for individual items $j$ are sampled from global vocabulary developmental parameters $\mu$ and $\sigma$ that characterize the mean and standard deviation of the intercepts and slopes for individual item trajectories.}
  \label{fig:bayesnet}
\end{figure}
%S2 represents the RSA speaker model defined by Eq. 1, which is used to predict the production responses dprod of each participant i, for each state s (number of hearts), for each utterance w, in each goal condition g. The RSA speaker model takes as input the literal meaning variables θ, which additionally are used to predict the literal meaning judgments dlit assuming a Bernoulli linking function. Additionally, the RSA model takes the speaker’s goal weights ω and intended presentational goal weight φ, which are inferred separately for each goal condition g. Finally, the RSA model uses two global free parameters: the cost of negation c and the speaker’s rationality parameter α.

>>>>

    \end{tabular}
  \end{center}
  \caption{Graphical model representing the Bayesian data analysis for \emph{Integration model} used for \emph{Explanation}. $L_1$ represents the RSA model defined by Eq. X, used to predict the Combination Data from Expt.~3 $d^{CM}$ for each unique age bin $i$, lexical item $j$, and alignment condition $k$. This RSA model takes as input a speaker optimality parameter $\alpha_i$ (which varies by age), a semantic knowledge parameter $\theta_{ij}$ (which varies by age and item), and a common ground sensitivity $\rho_i$ (which varies by age). Speaker optimality and semantic knowledge parameters are additionally constrained by the data from the Mutual Exclusivity task ($d^{ME}$; Expt.~1), via a related RSA model $L'_1$;  common ground sensitivity parameters $\rho_i$ are directly constrained by the data from the Common Ground experiment ($d^{CG}$; Expt.~2).  
  Each of these RSA parameters is sampled from a linear or logistic regression model that track developmental change (with intercepts and slopes given by $\beta_0$ and $\beta_1$, respectively). Additionally, the semantic knowledge parameters for individual items $j$ are sampled from global vocabulary developmental parameters $\mu$ and $\sigma$ that characterize the mean and standard deviation of the intercepts and slopes for individual item trajectories.}
  \label{fig:bayesnet}
\end{figure}
%S2 represents the RSA speaker model defined by Eq. 1, which is used to predict the production responses dprod of each participant i, for each state s (number of hearts), for each utterance w, in each goal condition g. The RSA speaker model takes as input the literal meaning variables θ, which additionally are used to predict the literal meaning judgments dlit assuming a Bernoulli linking function. Additionally, the RSA model takes the speaker’s goal weights ω and intended presentational goal weight φ, which are inferred separately for each goal condition g. Finally, the RSA model uses two global free parameters: the cost of negation c and the speaker’s rationality parameter α.

### 2020-04-14 paper/supplementary_material.Rmd
update supplement working in comments from MH

@@

Semantic knowledge captures the degree of certainty with which the naive listener is assumed to know the label for the familiar object. As a consequence, semantic knowledge differs between familiar objects. For objects whose labels are generally acquired earlier (e.g. carrot) semantic knowledge should generally be high whereas for others (e.g. pawn) semantic knowledge should generally be lower. However, semantic knowledge also varies with age such that older children are more likely to know the labels for more of the familiar objects compared to younger children. As a consequence, each familiar object has a unique developmental trajectory with respect to semantic knowledge. Technically, the object-specific parameters are estimated in the form of a hierarchical regression (mixed-effects) model; that is, each object's trajectory is estimated as a deviation from an overall trajectory of vocabulary development. This overall (vocabulary) trajectory represents the development of semantic knowledge that is independent of a particular familiar object--label pairing.

(MH: This paragraph is opaque)
Semantic knowledge enters the likelihood term of the model specified above. The likelihood depends not just on the parameter settings for semantic knowledge but also on the value of the parameter capturing expectations about speaker informativeness ($\alpha$ - see next section). As a consequence, these parameters are jointly estimated and co-vary with one another.

@@

A second locus of developmental change is a listener's expectations about speaker informativeness. In the context of the model, speaker informativeness corresponds to the degree with which the listener expects the speaker to choose the most informative of the available utterances. We assume that children at different ages could have different expectations about how rational or informative speakers are [see e.g. @bohn_tessler_merrick_frank_2019; @frank2014inferring; @yoon2019role]. As mentioned above, this parameter is jointly estimated with the parameters for semantic knowledge.

@@

We used the following prior distributions for model parameters. Intercept and slope for sensitivity to common ground: $unif(-2,2)$. For the speaker informativeness we used $unif(-3,3)$ for the intercept $unif(-0,4)$ for the slope. We restricted the slope to be positive because negative values for speaker informativeness are conceptually implausible. For semantic knowledge we used $unif(-3,3)$ for the intercept and $unif(0,2)$ for the slope, again because negative values of semantic knowledge have no meaningful interpretation. For the parameter capturing the variability the object specific trajectories around these overall parameters we used $unif(0,2)$ for the intercept and $unif(0,1)$ for the slope. Some choices regarding these prior distributions were made to ease model convergence. However, please note that all models considered used the same prior distributions.

@@

To estimate the parameters for semantic knowledge and speaker informativeness, we adapted the model described above to a situation in which both objects (novel and familiar) have equal prior probability (i.e., no common ground information). We used the data from Experiment 1 to then infer the parameters. That is, we inferred the intercepts and slopes for speaker informativeness and semantic knowledge that  generated RSA model predictions to match the responses generated in Experiment 1. To estimate the parameters representing sensitivity to common ground, we used a simple logistic regression to infer which combination of intercept and slope would generate predictions that corresponded to the average proportion of correct responses measured in Experiment 2.

@@

The first lesionsed model takes in speaker informativeness and common ground sensitivity as well as general semantic knowledge, but omits semantic knowledge that is specific to the familiar objects. We described above that the parameters for semantic knowledge are fitted via a hiearchical regression (mixed effects) model. In this model, there is an overall developmental trajectory for semantic knowledge (main effect) and then there is object-specific variation around this trajectory (random effects). The *no word knowledge* model takes in the overall trajectory for semantic knowledge but ignores object-specific variation. That is, the model assumes a listener whose mutual exclusivity inference does not vary depending on the particular familiar object but only depends on the average semantic knowledge.

@@

This model takes in object specific semantic knowledge and speaker informativeness but ignores common ground. Thus, the prior distribution over objects $P(r)$ in the model described above is uniform (e.g., [0.5, 0.5]). This corresponds to a listener who only focuses on the mutual exclusivity inference and ignores the common ground manipulation. As a consequence, the listener does not differentiate between the two common ground alignment conditions.

@@

The last lesion model only takes in the common ground sensitivity. This corresponds to a listener who only focuses on common ground and ignores the identity of the objects on the tables as well as any inferences their semantic knowledge of the familiar objects license. The model predictions therefore correspond to the prior distribution over objects $P(r)$.

@@

We compared the models mentioned above in two ways. On the one hand, we used correlations between model predictions and the data. For this analysis, we binned the model predictions and the data by age in years and familiar object. Figure \@ref(fig:fig6) visualizes the correlation between model predictions and the data for all models. The results shows a very high correlation between the predictions of the *integration model* and the data in all age groups suggesting that the model accurately captures the variation in the data. The correlation increased from 2- to 3-year-olds but then again dropped for 4-year-olds. The mis-match for the oldest age group is probably a consequence of the model making very extreme predictions for this group. [MH: Do we know why the model makes extreme predictions for the older age? Is it because they are at ceiling on the ME and/or common ground task, and our model posits no integration cost?] Correlations for the *integration model* were also higher compared to the other models considered.

We additionally compared models based on the marginal likelihood of the data under each model -- the likelihood of the data averaging over ("marginalizing over") the prior distribution on parameters; the pair-wise ratio of marginal likelihoods for two models is known as the Bayes Factor (see file `model_comparison.Rmd` in the associated online repository). Bayes Factors quantify the quality of predictions of a model, averaging over the possible values of the parameters of the models (weighted by their prior probability); by averaging over the prior distribution on parameters, Bayes Factor implicitly take into account model complexity because models with more parameters will tend to have a broader prior distribution over parameters, which in effect, can water down the possible better predictions that a model with more parameters can achieve (CITE: Lee & Wagenmakers 2014 Bayesian Cognitive Modeling). For this analysis, we treated age continuously. Table \@ref(tab:table5) lists the Bayes factors for the different model comparisons. The results show that the *integration model*, by far, outperformed all the other models. When comparing the lesion models among each other, we see that models including the mutual exclusivity inference make better predictions compared to the *no mutual exclusivity model*.

@@

The *integration* and *mixture* models have different implications for developmental change.
The *integration model* assumes that the process by which information is integrated remains constant across development. What changes is children's semantic knowledge, their expectations about speaker informativeness, and their sensitivity to common ground, but the way these information sources are combined remains the same across age. The *mixture model* alternatively posits that there is some bias or preference for one information source, and this bias could across development. This *developmental mixture model*  makes the same assumptions about developmental change in the individual information sources as the *integration model* but, in addition, assumes that the way that the the mutual exclusivity and the common ground inference are combined with one another changes over time. This model is structurally identical to the *mixture model* but the mixture parameter $\phi$ is a function of age.

@@

The *integration model* in this section differs from the integration model in the **Prediction Section** only in that the parameter distributions are now additionally informed by the data from Experiment 3. Figure \@ref(fig:figa1) - \@ref(fig:figa3) in the Appendix show how the parameter distributions differ between the prediction and the explanation version of the *integration* model. Semantic knowledge and speaker informativeness have similar posterior distributions when taking into account all the data compared to when estimating these parameters only based on experiment 1 and 2. In contrast, the intercept for common ground sensitivity is estimated to be larger and the slope shallower after taking into account Experiment 3 data. That is, our best guess for common ground sensitivity after taking into account the data from Experiment 3 is that children overall are more sensitive to the common ground manipulation and there is less developmental change. The code to run the model can be found in the associated online repository (file: xxxxxxx). Figure \@ref(fig:fig9) shows model predictions for the *integration model* in comparison to the data from experiment 3.

@@

Perhaps unsurprisingly, we see that updating the parameters for semantic knowledge, speaker informativeness and common ground sensitivity greatly improves the model fit (comparison: integration (explanation) vs integration (prediction)). We also see that the explanation *integration model* provides, by far, the best fit to the data compared to the two *mixture models*. Interestingly, the prediction *integration model* also had a better fit, even though its parameters were not updated based on the data from experiment 3. When comparing the two *mixture models* directly, we see that an age sensitive mixture parameter did not result in a substantially better fit.

This analysis shows that the inference and integration process described by the *ingetration model* accurately captures the data and also explains the integration process better compared to the additive *mixture models*. As a consequence, we may say that instead of being biased towards one type of inference, children are rationally integrating all the information sources available.

@@

Next, we tested the predictive power of this model. That is, we asked how well the model would predict the data of experiment 3, when only knowing the developmental trajectories for mutual exclusivity (based on experiment 1) and common ground (experiment 2). We found a very close alignment of the model predictions and the data across the entire age range. Furthermore, the *integration model* provided a better fit to the data compared to a number of lesion models, which selectively omitted one type of information. This suggests that children flexibly integrate all available information.

In the final section, we studied which process best explained children's information integration. We compared the *integration model* to a *mixture model* which assumed that children are biased towards one type of inference. We found that the *integration model* better explained the data compared to these alternative model. In sum, we found that children's integration of semantic and pragmatic information during word learning is best described as a form of rational social inference.

>>>>

```{r bda, include = T, fig.align = "center", fig.cap = "tbd."}
#knitr::include_graphics("../graphs/method_overview.png")
```

@@

Semantic knowledge captures the degree of certainty with which the naive listener is assumed to know the label for the familiar object. As a consequence, semantic knowledge differs between familiar objects. For objects whose labels are generally acquired earlier (e.g. carrot) semantic knowledge should generally be high whereas for others (e.g. pawn) semantic knowledge should generally be lower. However, semantic knowledge also varies with age such that older children are more likely to know the labels for more of the familiar objects compared to younger children. As a consequence, each familiar object has a unique developmental trajectory with respect to semantic knowledge. Technically, the object-specific parameters ($\theta_{ij}$) are estimated in the form of a hierarchical regression (mixed-effects) model; that is, each object's trajectory is estimated as a deviation from an overall trajectory of vocabulary development. This overall (vocabulary) trajectory represents the development of semantic knowledge that is independent of a particular familiar object--label pairing (see Figure \@ref(fig:bda)).

@@

A second locus of developmental change is a listener's expectations about speaker informativeness. In the context of the model, speaker informativeness corresponds to the degree with which the listener expects the speaker to choose the most informative of the available utterances. We assume that children at different ages could have different expectations about how rational or informative speakers are [see e.g. @bohn_tessler_merrick_frank_2019; @frank2014inferring; @yoon2019role].

@@

We used the following prior distributions for model parameters. Intercept and slope for sensitivity to common ground: $unif(-2,2)$. For the speaker informativeness we used $unif(-3,3)$ for the intercept $unif(-0,4)$ for the slope. We restricted the slope to be positive because negative values for speaker informativeness are conceptually implausible. For semantic knowledge we used $unif(-3,3)$ for the intercept and $unif(0,2)$ for the slope, because it is implausible to assume that semantic knowledge decreases with age. For the parameter capturing the variability the object specific trajectories around these overall parameters we used $unif(0,2)$ for the intercept and $unif(0,1)$ for the slope. Some choices regarding these prior distributions were made to ease model convergence. However, please note that all models considered used the same prior distributions.

@@

To estimate the parameters for semantic knowledge and speaker informativeness, we adapted the model described above to a situation in which both objects (novel and familiar) have equal prior probability (i.e., no common ground information). We used the data from Experiment 1 to then infer the parameters. That is, we inferred the intercepts and slopes for speaker informativeness (linear regression) and semantic knowledge (logistic regression) that generated RSA model predictions to match the responses generated in Experiment 1. To estimate the parameters representing sensitivity to common ground, we used a simple logistic regression to infer which combination of intercept and slope would generate predictions that corresponded to the average proportion of correct responses measured in Experiment 2.

@@

The first lesionsed model takes in speaker informativeness and common ground sensitivity as well as general semantic knowledge, but omits semantic knowledge that is specific to the familiar objects. We described above that the parameters for semantic knowledge are fitted via a hierarchical regression (mixed effects) model. In this model, there is an overall developmental trajectory for semantic knowledge (main effect) and then there is object-specific variation around this trajectory (random effects). The *no word knowledge* model takes in the overall trajectory for semantic knowledge but ignores object-specific variation. That is, the model assumes a listener whose mutual exclusivity inference does not vary depending on the particular familiar object but only depends on the average semantic knowledge.

@@

This model takes in object specific semantic knowledge and speaker informativeness but ignores common ground. Thus, the prior distribution over objects in the model described above is uniform (e.g., [0.5, 0.5]). This corresponds to a listener who only focuses on the mutual exclusivity inference and ignores the common ground manipulation. As a consequence, the listener does not differentiate between the two common ground alignment conditions.

@@

The last lesion model only takes in the common ground sensitivity. This corresponds to a listener who only focuses on common ground and ignores the identity of the objects on the tables as well as any inferences their semantic knowledge of the familiar objects license. The model predictions therefore correspond to the prior distribution over objects $P(r \mid \rho_i)$.

@@

We compared the models mentioned above in two ways. On the one hand, we used correlations between model predictions and the data. For this analysis, we binned the model predictions and the data by age in years and familiar object. Figure \@ref(fig:fig6) visualizes the correlation between model predictions and the data for all models. The results shows a very high correlation between the predictions of the *integration model* and the data in all age groups suggesting that the model accurately captures the variation in the data. Correlations for the *integration model* were also higher compared to the other models considered. The correlation increased from 2- to 3-year-olds but then again dropped for 4-year-olds. The mis-match for the oldest age group is probably a consequence of the model making very extreme predictions in the congruent condition. This results from the fact that 4-year-olds show very high performance in the mututal exclusivity as well as the common ground task. When combined in the model, the two inferences amplify one another because we do not assume that there is any cost of integration. We made this choice because the parameter representing integration cost would have to be estimated based on the integration data itslef.

We additionally compared models based on the marginal likelihood of the data under each model -- the likelihood of the data averaging over ("marginalizing over") the prior distribution on parameters; the pair-wise ratio of marginal likelihoods for two models is known as the Bayes Factor (see file `model_comparison.Rmd` in the associated online repository). Bayes Factors quantify the quality of predictions of a model, averaging over the possible values of the parameters of the models (weighted by their prior probability); by averaging over the prior distribution on parameters, Bayes Factor implicitly take into account model complexity because models with more parameters will tend to have a broader prior distribution over parameters, which in effect, can water down the possible better predictions that a model with more parameters can achieve [@lee2014bayesian]. For this analysis, we treated age continuously. Table \@ref(tab:table5) lists the Bayes factors for the different model comparisons. The results show that the *integration model*, by far, outperformed all the other models. When comparing the lesion models among each other, we see that models including the mutual exclusivity inference make better predictions compared to the *no mutual exclusivity model*.

@@

The *integration* and *mixture* models have potentially different implications for developmental change.
The *integration model* assumes that the process by which information is integrated remains constant across development. What changes is children's semantic knowledge, their expectations about speaker informativeness, and their sensitivity to common ground, but the way these information sources are combined remains the same across age. The *mixture model* alternatively posits that there is some bias or preference for one information source, and this bias could change across development. This *developmental mixture model*  makes the same assumptions about developmental change in the individual information sources as the *integration model* but, in addition, assumes that the way that the the mutual exclusivity and the common ground inference are combined with one another changes over time. This model is structurally identical to the *mixture model* but the mixture parameter $\phi$ is a function of age (i.e. represented by an intercept and a slope).

@@

The *integration model* in this section differs from the integration model in the **Prediction Section** only in that the parameter distributions are now additionally informed by the data from Experiment 3. Figure \@ref(fig:figa1) - \@ref(fig:figa3) in the Appendix show how the parameter distributions differ between the prediction and the explanation version of the *integration* model. Semantic knowledge and speaker informativeness have similar posterior distributions when taking into account all the data compared to when estimating these parameters only based on experiment 1 and 2. In contrast, the intercept for common ground sensitivity is estimated to be larger and the slope shallower after taking into account Experiment 3 data. That is, our best guess for common ground sensitivity after taking into account the data from Experiment 3 is that younger children are more sensitive to the common ground manipulation and there is less developmental change. The code to run the model can be found in the associated online repository (file: xxxxxxx). Figure \@ref(fig:fig9) shows model predictions for the *integration model* in comparison to the data from experiment 3.

@@

Perhaps unsurprisingly, we see that further constraining the parameters for semantic knowledge, speaker informativeness and common ground sensitivity by the data from experiment 3 greatly improves the model fit (comparison: integration (explanation) vs integration (prediction)). We also see that the explanation *integration model* provides, by far, the best fit to the data compared to the two *mixture models*. Interestingly, the prediction *integration model* also had a better fit, even though its parameters were not constrained by the data from experiment 3. When comparing the two *mixture models* directly, we see that an age sensitive mixture parameter did not result in a substantially better fit.

This analysis shows that the inference and integration processes described by the *ingetration model* accurately capture the data and also explain information integration better compared to the additive *mixture models*. As a consequence, we may say that instead of being biased towards one type of inference, children are rationally integrating all the information sources available.

@@

Next, we tested the predictive power of this model. That is, we asked how well the model would predict the data of experiment 3, when only knowing the developmental trajectories for mutual exclusivity (based on experiment 1) and common ground (experiment 2). We found a very close alignment of the model predictions and the data across the entire age range. Furthermore, the *integration model* provided a better fit to the data compared to a number of lesioned models, which selectively omitted one type of information. This suggests that children flexibly integrate all available information.

In the final section, we studied which process best explained children's information integration. We compared the *integration model* to a *mixture model* which assumed that children are biased towards one type of inference. We found that the *integration model* better explained the data compared to these alternative model. In sum, we found that children's integration of semantic and pragmatic information during word learning is best described as a form of Bayesian social inference.
